Population Study

### 2021-04-11 submissiondraftIMT.Rmd
Included more about data, model, results.

>>>>

Because COVID cases can vary largely by state due to the size of the state, we were careful in how we constructed the $\mbox{COVID}_{s,t}$ covariate. We used the log ratio of new cases in sequential weeks as the COVID covariate:
$$\mbox{COVID}_{s,t} = \log\left(\frac{\text{New cases in state $s$ and week $t$}}{\text{New cases in state $s$ and week $t-1$}}\right).$$

This value is appealing for three reasons. First, it places all states on the same scale by only considering the ratio of subsequent weeks. Second, by using a ratio of new covid cases it handles the exponential growth nature of the pandemic. And third, it works well with an autoregressive model. A positive value means the pandemic is currently accelerating in a state, possibly resulting in increasing anxiety relative to week $t-1$. Similarly, a negative value means the pandemic is slowing down, possibly resulting in decreasing anxiety relative to week $t-1$.

### Results

```{r getmodeldata, include=FALSE}
library(plm)

# Read in anxiety data, covid data (whole US) and covid data (by state)
anxiety <- read.csv("anx_rate_ext.csv")
covid.us <- read.csv("covid-data/weekly-us.csv")
covid <- read.csv("covid-data/weekly-states.csv")

# Create logitudinal long form data frame
paneldata <- data.frame(
  State = covid$State,
  Date = covid$EndofWeek,
  CasesPerDay = covid$Cases,
  LogCasesRatio = covid$LogRatio
)
for (i in 1:nrow(paneldata)) {
  s <- paneldata[i, "State"]
  d <- paneldata[i, "Date"]

  if (d %in% anxiety$date_end) {
    paneldata[i,"anx"] <- anxiety[anxiety$date_end == d, s]
  }
}

paneldata <- paneldata[paneldata$Date %in% anxiety$date_end,] # Only dates with an anxiety measurement
```

First, we have the output from the fixed effects model:
```{r runfixedmodel}
# Model: panel model by state with anxiety as response regressed on covid
fit.states.fixed <- plm(anx ~ LogCasesRatio + lag(anx), paneldata, 
                        index=c("State", "Date"), model="within")
summary(fit.states.fixed)
```

We can see that both the lagged anxiety variable and LogCasesRatio are significant predictors. The value of the coefficient with $\mbox{COVID}_{s,t}$ shows that if new cases were to double from one week to the next, we expect our response anxiety odds to increase by `r round(fit.states.fixed$coefficients[1] * log(2), digits=4)`.

The `plm` package also allows us to run a Hausman hypothesis test to see whether a fixed effects or random effects model is a better fit for our data.

```{r runrandommodel}
fit.states.random <- plm(anx ~ LogCasesRatio + lag(anx), paneldata, 
                         index=c("State", "Date"), model="random")
phtest(fit.states.fixed, fit.states.random)
```

The small p-value indicates that the null hypothesis of a random effects model should be rejected, and so our results from the fixed effects model are a better fit.
